<!-- %\VignetteEngine{knitr::knitr} %\VignetteIndexEntry{Using rprime} -->

```{r, echo = FALSE, message = FALSE}
library("rprime")
library("knitr")
opts_chunk$set(
  comment = "#>",
  error = FALSE,
  tidy = FALSE)
```

## rprime

Eprime, or more officially E-Prime® 2.0, is a set of programs by [Psychology Software
Tools, Inc.](http://www.pstnet.com/) used for running psychological
experiments. Data from experiment sessions are saved into two files: 1) a
proprietary [binary `.edat2`
file](http://www.pstnet.com/support/kb.asp?TopicID=3149), and 2) a plain-text
`.txt` file. The rprime package provides a set of functions for parsing these
`.txt` files inside R.

**Disclaimer**: I write this documentation as someone who has very little time
creating and designing experiments in Eprime but has spent a lot of time
parsing Eprime text files in R. Therefore, this vignette is not documentation
for Eprime software and its data format. Instead, this vignette just outlines
this package's approach to extracting data from Eprime text files.


### Background

An Eprime `txt` file is a series of **frames**. Each frame is a list of
key-value pairs. Frames are bracketed by special start and stop lines.

#### Headers

Each file begins with a special **Header** frame, which looks something like
this:

```
*** Header Start ***
VersionPersist: 1
LevelName: Session
LevelName: Block
LevelName: Trial
LevelName: SubTrial
LevelName: LogLevel5
LevelName: LogLevel6
LevelName: LogLevel7
LevelName: LogLevel8
LevelName: LogLevel9
LevelName: LogLevel10
Experiment: SAE_MinimalPairDiscrim
SessionDate: 07-10-2013
SessionTime: 10:46:22
SessionTimeUtc: 3:46:22 PM
Dialect: Yes
Subject: 001
ExpTyp: L
Age: 00
Gender: X
Session: 1
RandomSeed: -261296366
Group: 1
Display.RefreshRate: 60.003
Clock.Information: <?xml version="1.0"?>\n<Clock xmlns:dt="urn:schemas-microsoft-com:datatypes"><Description dt:dt="string">E-Prime Primary Realtime Clock</Description><StartTime><Timestamp dt:dt="int">0</Timestamp><DateUtc dt:dt="string">2013-07-10T15:46:22Z</DateUtc></StartTime><FrequencyChanges><FrequencyChange><Frequency dt:dt="r8">3020527</Frequency><Timestamp dt:dt="r8">364810315607</Timestamp><Current dt:dt="r8">0</Current><DateUtc dt:dt="string">2013-07-10T15:46:22Z</DateUtc></FrequencyChange></FrequencyChanges></Clock>\n
*** Header End ***
```

The header is set off by the brackets `*** Header Start***` and `*** Header
End***`. The fields in the header describe the basic settings of the
experiment.

#### LogFrames

The other frames in the text file are **LogFrames**. These record data about
individual the procedures or trials are executed during the experiment. The
following lines immediately follow the header from the previous example:

```
	Level: 2
	*** LogFrame Start ***
	Procedure: FamTask
	item1: bear
	item2: chair
	CorrectResponse: bear
	ImageSide: Left
	Duration: 885
	Familiarization: 1
	FamInforcer: 1
	ReinforcerImage: Bicycle1
	Familiarization.Cycle: 1
	Familiarization.Sample: 1
	Running: Familiarization
	FamTarget.RESP:
	Correct: True
	*** LogFrame End ***
```

Trials and procedures can be **nested** inside of higher-order sections of the
experiment. For example, practice trials may be nested inside of a practice
block of an experiment. The level of nesting in the above frame is indicated by
the number of tabs before each line as well as the `Level: ` line. The only
Level 1 frames in an experiment appear to the header frame and the final frame
off an experiment, which largely duplicated the header frame.

Some **special fields** appear in every log-frame. These fields are
`Procedure`, `Running`, `[Running].Sample` and `[Running].Cycle` where
`[Running]` is the value of the running field. In the previous example, these
were:

```
	Procedure: FamTask
	Familiarization.Cycle: 1
	Familiarization.Sample: 1
	Running: Familiarization
```

When trials are presented in a random order, the `[Running].Sample` field records the sequential trial number. 


### Basic strategy 


#### File reading

First, the lines from the text file are read into R and stored into a character vector. Here are the first few lines from the example file:

```{r}
library("rprime")
exp_lines <- read_eprime("data/MINP_001L00XS1.txt")
head(exp_lines)
```

If the file is not an Eprime `.txt` file, a **dummy header** is created so that the 
file may be treated like any other like Eprime. The user is warned as this happens.

```{r}
bad_lines <- read_eprime("data/not_an_eprime_file.txt")
head(bad_lines)
```

I chose to use a dummy header instead of raising an error so that code which loaded 
multiple files at once would not fail outright when it encountered a bad file.


#### Chunking

The basic strategy for paring the Eprime file: 

1. Read in the Eprime file
2. Creating EprimeFrame objects
    a. Extract a frame by pulling out lines of text that fall between a pair of 
       bracketing lines.
    b. Create key-value pairs inside the frame by splitting text lines at `:`s 
    c. Add some helpful metadata inside each frame.
    d. Bundle up the key-value pairs into an R `list` object with an added 
       `Eprime.Frame` class.
3. Storing the frames from an experiment together inside of a `FrameList` object.

```{r}


exp_chunk <- exp_lines[28:44]
EprimeFrame(exp_chunk)
FrameList(exp_chunk)
```


The basic strategy for

```{r}
FrameList(exp_lines)[[2]]
rprime::
EprimeFrame(exp_lines)
```


