---
title: "Quick start demonstration"
author: "Tristan Mahr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick start demonstration}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
library("rprime")
library("knitr")
opts_chunk$set(
  comment = "#>",
  error = FALSE,
  tidy = FALSE,
  collapse = TRUE)
```

**rprime** is an R package for parsing `.txt` generated by E-Prime, a program 
for running psychological experiments. 


Overview
-------------------------------------------------------------------------------

The main workflow for working with rprime involves:

1. `read_eprime`: reliably read in data from an Eprime log (`.txt`) file.
2. `FrameList`: extract the text in each `"LogFrame"` in the file, storing each 
   log-frame as an R list.
3. `preview_levels`, `preview_eprime`: explore the structure of the parsed 
   data.
4. `keep_levels`, `drop_levels`, `filter_in`, `filter_out`: select and filter 
   particular levels from the txt-file.
5.  `to_data_frame`: make a data-frame from the parsed data.


Installation
-------------------------------------------------------------------------------

To get the current, released version from CRAN:

```{r, eval = FALSE}
install.packages("rprime")
```


Examples
-------------------------------------------------------------------------------

### Getting data into R

Load the file with `read_eprime` and parse its contents with `FrameList`.

```{r}
library("rprime")
# Read in an Eprime text file
experiment_lines <- read_eprime("data/SAILS/SAILS_001X00XS1.txt")

# Extract and parse the log-frames from the file
experiment_data <- FrameList(experiment_lines)
```

### Exploring

In the text file, frames were distinguished by the procedure they are running as 
well as the their level of nesting. Get an overview of the different types of 
frames with `preview_levels`.

```{r}
# There are six different kinds of frames in this file
preview_levels(experiment_data)
```

Get a preview of the data in each kind of frame with `preview_frames`.

```{r}
preview_frames(experiment_data)
```

`preview_eprime` (not demonstrated here) does both kinds of previews (levels 
and frames).

### Filtering

Use `keep_levels` and `drop_levels` to filter frames according to nesting level.

```{r}
# Filter (out) by depth of nesting
not_level_1 <- drop_levels(experiment_data, 1)
preview_levels(not_level_1)

# Filter (in) by depth of nesting
just_level_3 <- keep_levels(experiment_data, 3)
preview_levels(just_level_3)
```

Use `filter_in` and `filter_out` to filter frames using attribute values. Use 
repeated filtering statements to drill down into the list of frames.

```{r}
# Filter (out) by attribute values
no_header <- filter_out(experiment_data, "Running", values = "Header")
preview_levels(no_header)

# Filter (in) by attribute values
not_practice <- filter_in(experiment_data, "Running", "TrialLists")
preview_levels(not_practice)

# Drill down further into the trials by filtering again
sue_trials <- filter_in(not_practice, "Module", "SUE")
preview_eprime(sue_trials)
```

### Exporting

Convert to a dataframe with `to_dataframe`. Attribute names in the log-frames 
become column names in the dataframe.

```{r}
# Export to dataframe
sue_trials_df <- to_data_frame(sue_trials)
str(sue_trials_df, vec.len = 2)

# Don't need every column
columns_to_keep <- c("Eprime.Basename", "Module", "Sample", 
                     "Correct", "Response")
sue_trials_df <- sue_trials_df[columns_to_keep]
head(sue_trials_df)
```
