
<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Using rprime}
-->

```{r, echo = FALSE, message = FALSE}
library("rprime")
library("knitr")
opts_chunk$set(
  comment = "#>",
  error = FALSE,
  tidy = FALSE)
```

## Quick Start Demonstration

```{r}
# Read in an Eprime text file
experiment_lines <- read_eprime("data/SAILS/SAILS_001X00XS1.txt")

# Extract and parse the log-frames from the file
experiment_data <- FrameList(experiment_lines)

# Run-down of different log-frames in the FrameList
preview_levels(experiment_data)

# Filter (out) by depth of nesting
not_level_1 <- drop_levels(experiment_data, 1)
preview_levels(not_level_1)

# Filter (in) by depth of nesting
just_level_3 <- keep_levels(experiment_data, 3)

# Preview each kind of log-frame in the FrameList
preview_eprime(just_level_3)

# Filter (out) by attribute values
no_header <- filter_out(experiment_data, "Running", values = "Header")
preview_levels(no_header)

# Filter (in) by attribute values
not_practice <- filter_in(experiment_data, "Running", "TrialLists")
preview_eprime(not_practice)

# Drill down into a FrameList
sue_trials <- filter_in(not_practice, "Module", "SUE")
preview_eprime(sue_trials)

# Export to dataframe
columns_to_keep <- c("Eprime.Basename", "Module", "Sample", "Correct", "Response")
sue_trials_df <- to_data_frame(sue_trials)[columns_to_keep]
head(sue_trials_df)
```


## Working with multiple files

Suppose we want to load all the Eprime files in a directory and combine the results in data-frame. 

My strategy in this scenario is to figure out what I need to do for a single file and then wrap those steps in a function that takes a filepath to a txt file and returns a data-frame. After some exploration and interactive programming, I come up with the following function.

```{r}
library("plyr")
reduce_sails <- function(sails_path) {
  sails_lines <- read_eprime(sails_path)
  sails_frames <- FrameList(sails_lines)
  
  # Trials occur at level 3
  sails_frames <- keep_levels(sails_frames, 3)
  sails <- to_data_frame(sails_frames)
  
  # Tidy up
  to_pick <- c("Eprime.Basename", "Running", "Module", "Sound", 
               "Sample", "Correct", "Response")
  sails <- sails[to_pick]
  running_map <- c(TrialLists = "Trial", PracticeBlock = "Practice")
  sails$Running <- ifelse(sails$Running == "PracticeBlock", "Practice", "Trial")
  
  # Renumber trials in the practice and experimental blocks separately.
  # Numerically code correct response.
  sails <- ddply(sails, .(Running), mutate, 
                 TrialNumber = seq(from = 1, to = length(Running)),
                 CorrectResponse = ifelse(Correct == Response, 1, 0))
  sails$Sample <- NULL
  
  # Optionally, one might save the processed file via: 
  # csv <- paste0(file_path_sans_ext(sails_path), ".csv")
  # write.csv(sails, csv, row.names = FALSE)
  sails
}
```

Here's a preview of what the function returns when given a filepath.

```{r}
head(reduce_sails("data/SAILS/SAILS_001X00XS1.txt"))
```

Now that the function works on one file, I can use `ldply` to apply the function to several files, returning results in a single dataframe.

```{r}
sails_paths <- list.files("data/SAILS/", pattern = ".txt", full.names = TRUE)
sails_paths
ensemble <- ldply(sails_paths, reduce_sails)
```

Finally, with all of the subjects' data contained in a single data-frame, I can use `ddply` plus `summarise` and compute summary scores at different levels of aggregation within each subject. 

```{r}
# Score trials within subjects
overall <- ddply(ensemble, .(Eprime.Basename, Running), summarise, 
                 Score = sum(CorrectResponse),
                 PropCorrect = Score / length(CorrectResponse))
overall

# Score modules within subjects
modules <- ddply(ensemble, .(Eprime.Basename, Running, Module), summarise, 
                 Score = sum(CorrectResponse),
                 PropCorrect = Score / length(CorrectResponse))
modules

# Drop practice values, then print data in wide format
library("reshape2")
dcast(modules, Eprime.Basename ~ Module, subset = .(Running == "Trial"), 
      fun.aggregate = mean, value.var = "PropCorrect")
```
